---
title: "Preparing Data for Regions"
output: 
  html_document:
    theme: yeti
---

Run in R 3.5.2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(mapview))
suppressPackageStartupMessages(library(leaflet))
suppressPackageStartupMessages(library(miniUI))
suppressPackageStartupMessages(library(DT))

source('snapFunctions.R') # snapping functions
source('snapOrganizationFunctions_v2.R') # functions to do stuff with snapping functions
```

This document walks users through the requisite data preprocessing steps to prepare data for the 2020 IR Rivers and Streams Assessment and 2020 IR Lakes Assessment decision support applications. All initial steps are tested on 2016 and 2018 IR data and thus need to be rerun for 2020 data when those datasets become available.

This document demonstrates the necessary steps to get any regional office up to speed with prerequisite data organization/processing steps to run the 2020 IR. 

## Input data

#### Conventionals
Bring in Roger's conventionals dataset. Make it a csv immediately to speed up rendering in app. This is the final data pull from Roger for the 2018 IR. As of today, the 2020 final IR pull is not available. **Subsitute 2020 data in when available.**

```{r conventionals2018, echo=F}
#conventionals <- read_excel('workingDatasets/CONVENTIONALS_20171010.xlsx',sheet = 'CONVENTIONALS')
#conventionals$FDT_DATE_TIME2 <- as.POSIXct(conventionals$FDT_DATE_TIME, format="%m/%d/%y %H:%M")
#write.csv(conventionals, 'workingDatasets/CONVENTIONALS_20171010.csv', row.names=F)

conventionals <- suppressWarnings(suppressMessages(read_csv('workingDatasets/CONVENTIONALS_20171010.csv')))
#View(conventionals) # to look at dataset use this command
```

Before we can do any spatial subsetting work, we need to make a dataset of all UNIQUE StationID's for the regional offices to work with. 

```{r conventionalsDistinct, echo=FALSE}
conventionals_D <- distinct(conventionals, FDT_STA_ID, .keep_all = T) %>%
  select(FDT_STA_ID:FDT_SPG_CODE, STA_LV2_CODE:STA_CBP_NAME) %>%# drop data to avoid any confusion
  st_as_sf(coords = c("Longitude", "Latitude"),  # make spatial layer using these columns
           remove = F, # don't remove these lat/lon cols from df
           crs = 4326) # add coordinate reference system, needs to be geographic for now bc entering lat/lng, 
rm(conventionals) # remove full conventionals dataset to save memory

#View(conventionals_D) # to look at dataset use this command
```


Below is the table output. For this cycle, there are `r nrow(conventionals_D)` unique stations sampled.

```{r conventionalsDistinctTable, echo=FALSE}
DT::datatable(conventionals_D, escape=F, rownames = F, options = list(scrollX = TRUE))
#View(conventionals_D)
```


#### Statewide Assessment Layer

Because we cannot trust the Deq_Region or STA_REC_CODE columns from conventionals to give us the stations each region needs to assess, we must do a spatial subset of all unique stations in an IR window (conventionals_D) by the Statewide Assessment Layer (filtered to our specified assessment region).

First, read in the statewide assessment layer.

```{r Statewide Assessment Layer, echo=FALSE, message=FALSE, warning=FALSE, results = 'hide'}
assessmentLayer <- st_read('GIS/AssessmentRegions_VA84_basins.shp') %>%
  st_transform( st_crs(4326)) # transform to WQS84 for spatial intersection
```

Now we are going to subset all unique stations from conventionals to a specific region. **This requires minimal user input**
 
Below are the regional codes you need to choose from 
```{r regionalCodes}
unique(assessmentLayer$ASSESS_REG)
```

Insert one of those unique codes in quotes into the chunk below. The commented out script serves as an example for you.

```{r DEQ region selection}
assessmentLayerSelection <- filter(assessmentLayer, ASSESS_REG == 'BRRO')
#assessmentLayerSelection <- filter(assessmentLayer, ASSESS_REG ==    )
#assessmentLayerSelection <- filter(assessmentLayer, ASSESS_REG == 'BRRO')
```

Now we can subset unique conventionals sites that fall into the assessment region we want.

```{r conventionals_D_Region}
conventionals_D_Region <- st_join(conventionals_D, assessmentLayerSelection, join = st_intersects) %>%
  filter(!is.na(VAHU6)) %>% # only keep rows that joined
  mutate(sameVAHU6 = ifelse(VAHU6 == Huc6_Vahu6, T, F)) %>% # double check the VAHU6 watershed matches from CEDS to actual spatial layer
  dplyr::select(FDT_STA_ID:STA_CBP_NAME, VAHU6, sameVAHU6) %>% # just get columns we are interested in
  rename('Basin' =  'Basin.x') %>% # rename column to original name pre join
  dplyr::select(FDT_STA_ID:STA_CBP_NAME) # get rid of the extra VAHU6 columns

rm(conventionals_D); rm(assessmentLayer) # clean up workspace

cat(paste('Below are the ',nrow(conventionals_D_Region),' unique stations in the ', unique(assessmentLayerSelection$ASSESS_REG),' region.', sep=''))
```


```{r DistinctSitesInRegionTable, echo=FALSE}
DT::datatable(conventionals_D_Region, escape=F, rownames = F, options = list(scrollX = TRUE))
#View(conventionals_D_Region)
```


#### Last Cycle Estuary Assessment Layer

First thing we want to do now is filter out any sites that fall into the estuary AUs. These stations will have to be assessed through other means. **Note: you can download the table below for reference by running the commented script. These are all the stations within the selected assessment region that fall into an estuary AU based on the 2018 AU layers.**

The below chunk:
1) Reads in the 2018 estuary AU layer
2) Finds any sites that fall into an estuary AU polygon
  - see optional code to write those sites to a separate csv for future reference
3) Removes any estuarine sites from the data frame of unique conventionals sites that need AU and WQS information.

```{r Filter Out Estuary Statewide, echo=F, results = 'hide'}
estuaryAUs <- st_read('GIS/draft2018IR_AUs/va_2018_aus_estuarine.shp') %>%
   st_transform( st_crs(4326)) # transform to WQS84 for spatial intersection


conventionals_D_Region_Estuary <- st_join(conventionals_D_Region, estuaryAUs, join = st_intersects) %>%
      filter(!is.na(OBJECTID))

#write.csv(conventionals_D_Region_Estuary, 'conventionals_D_Region_Estuary.csv', row.names = F) 
# you can change the file location like so:
#write.csv(conventionals_D_Region_Estuary, 'C:/Assessment/2020/dataProcessing/conventionals_D_Region_Estuary.csv', row.names = F) 

conventionals_D_Region_NoEstuary <- filter(conventionals_D_Region, !(FDT_STA_ID %in% conventionals_D_Region_Estuary$FDT_STA_ID))
rm(estuaryAUs) # remove from memory
```

Below is the table of sites that are estuarine.

```{r estuarySitesTable, echo=F}
DT::datatable(conventionals_D_Region_Estuary, escape=F, rownames = F, options = list(scrollX = TRUE))
#View(conventionals_D_Region_Estuary)
```

```{r inlineText2,echo=F}
cat(paste('Removing the above sites, the unique conventionals sites have gone from ', nrow(conventionals_D_Region),' to ',nrow(conventionals_D_Region_NoEstuary),'. The next step is to find the lake/reservoir stations and link them to the appropriate AU and WQS for analyses in the Lake Assessment App.', sep=''))
```

### Assessment Unit Organization

#### Last Cycle Stations Table

The next step uses the 2018 IR station table as a starting point for linking the stations from the 2020 IR window to the appropriate AU. This is important to do prior to doing any spatial subsetting or snapping because most of the stations from the previous cycle already have QAed assessment unit information readily available and organized. It is computationally less expensive to save any necessary spatial subsetting or snapping to lake/reservoir polygons or stream segments until all joins that can be done with table methods are first completed.

```{r previousCycleStationTable}
## This is JUST FOR BRRO RIGHT NOW BECAUSE THERE IS NOT FINAL 2018 STATIONS TABLE
## I NEED TO REWRITE TO GET TO SOMETHING LIKE X:\2016_Assessment\GIS_2016\MonitoringStations\ir_mon_stations_2016_FINAL.mdb


################################### Only this way for now because dont have one table I can pull this data from ##########################
# From Mary (Rivers)
stationTable <- read_excel('data/Emma_Stations2018IR_Draft_Dec.xlsx')[,1:48]

stationTable1 <- stationTable %>%
  group_by(STATION_ID) %>%
  mutate(extra= n()) %>%
  select(STATION_ID, extra, everything())

stationTable2 <- distinct(stationTable, STATION_ID, ID305B_1, ID305B_2, ID305B_3, .keep_all = T)%>%
  group_by(STATION_ID) %>%
  mutate(extra= n()) %>%
  select(STATION_ID, extra, everything()) %>% # still extras so just take distinct STATION_ID 
  distinct(STATION_ID,.keep_all = T) %>%
  select(-extra) %>%
  rename("COMMENTS" = "Notes") %>%
  select(ID305B_1,ID305B_2,ID305B_3,DEPTH,STATION_ID,REGION,everything()) %>%
  as_tibble()
rm(stationTable1)

#From Paula (lakes)
stationTableLake <- read_excel('data/Paula_tbl_ir_mon_stations_BRRO_2018_JRH_v1 (2).xlsx') %>%
  dplyr::select(ID305B_1:COMMENTS)
stationTableLake$SED_MET_VIO <- as.numeric(stationTableLake$SED_MET_VIO)

# Now combine and get rid of duplicate entries
previousCycleStationTable <- bind_rows(stationTable2,stationTableLake) %>%
  distinct(stationTable, STATION_ID, ID305B_1, ID305B_2, ID305B_3, .keep_all = T) %>%
  group_by(STATION_ID) %>%
  mutate(extra= n()) %>%
  select(extra, everything())

# special steps because still duplicates with VAC
dupes <- filter(previousCycleStationTable, extra>1) %>%
  distinct(STATION_ID,.keep_all = T) %>%
  select(-extra)


previousCycleStationTable <- filter(previousCycleStationTable, extra==1) %>%
  select(-extra) %>%
    bind_rows(dupes) %>%
  filter(!is.na(ID305B_1))

rm(stationTable2);rm(stationTableLake);rm(stationTable); rm(dupes)
##############################################################################################################


# Join unique sites from current cycle to last cycle's station table to see what already has AU info
conventionals_D_Region_AU <- mutate(conventionals_D_Region_NoEstuary, STATION_ID = FDT_STA_ID) %>% # make joining column
    left_join(previousCycleStationTable, by='STATION_ID') # join to get ID305B info

# Only work with sites that don't already have AU info
conventionals_D_Region_NoAU <- filter(conventionals_D_Region_AU, is.na(ID305B_1))
#View(conventionals_D_Region_NoAU)  

# Keep the stations with AU information to the side for later use
conventionals_D_Region_AU <- filter(conventionals_D_Region_AU, !is.na(ID305B_1))
#View(conventionals_D_Region_AU)

rm(conventionals_D_Region_NoEstuary) # clean up workspace
```


The table below illustrates all the  unique conventionals sites that matched stations from 2018 IR station table by STATION_ID/FDT_STA_ID.
```{r conventionals_D_Region_AU, echo=F}
DT::datatable(conventionals_D_Region_AU, escape=F, rownames = F, options = list(scrollX = TRUE))
#View(conventionals_D_Region_Estuary)
```

The table below illustrates all the unique conventionals sites that did not match any stations from 2018 IR station table by STATION_ID/FDT_STA_ID.
```{r conventionals_D_Region_NoAU, echo=F}
DT::datatable(conventionals_D_Region_NoAU, escape=F, rownames = F, options = list(scrollX = TRUE))
#View(conventionals_D_Region_Estuary)
```

#### Last Cycle Reservoir Assessment Layer 

```{r inlineText3, echo=F}
cat(paste('Working with the ', nrow(conventionals_D_Region_NoAU),' sites that did not connect to a STATION_ID/FDT_STA_ID from last cycle, we will first do a spatial join to the reservoir/lakes AUs to see if any sites are lake sites before using the more computationally intensive stream snapping process.', sep=''))
```

```{r reservoirStatewide, echo=F, results = 'hide'}
reservoirAUs <- st_read('GIS/draft2018IR_AUs/va_2018_aus_reservoir.shp') %>%
   st_transform( st_crs(4326)) # transform to WQS84 for spatial intersection

# find lake AUs, if Any
conventionals_D_Region_Lakes <- st_join(conventionals_D_Region_NoAU, reservoirAUs, join = st_intersects) %>%
  filter(!is.na(OBJECTID)) %>% # only keep rows where join successful
  mutate(ID305B_1 = ID305B) %>% # rename ID305B_1 field by the joined reservoir ID305B
  dplyr::select(names(conventionals_D_Region_AU))

# add lake AU sites to conventionals_D_Region_AU
conventionals_D_Region_AU <- rbind(conventionals_D_Region_AU, conventionals_D_Region_Lakes)

# and remove lake AU sites from the sites without AUs
conventionals_D_Region_NoAU <- filter(conventionals_D_Region_NoAU, !(FDT_STA_ID %in% conventionals_D_Region_Lakes$FDT_STA_ID))
rm(reservoirAUs); rm(conventionals_D_Region_Lakes) # remove from memory
```

#### Last Cycle Riverine Assessment Layer

```{r inlineText4, echo=F}
cat(paste('The last step for AU organization after we have removed any estuary sites and dealt with all sites that are in lake AUs is to automatically snap the', nrow(conventionals_D_Region_NoAU), 'remaining sites to riverine AUs.',sep=' '))
```

A custom, automated snapping tool is run below. The function buffers each site using the user defined sequence (bufferDistances). You do not need to change the bufferDistances variable, but commented out code demonstrates how you can if you want to increase/decrease the maximum buffer distance or breaks. 

The function first buffers the site the lowest input buffer distance, then it intersects that buffer with the input assessment unit layer. If no segments are retrieved in the given buffer distance, then the next highest buffer is tried and so on until the maximum buffer distance is tested. If one or more segments are captured in a given buffer, then the buffering process stops, returns the snapped stream segment information, and moves to the next site. If more that one site is captured in a single buffer, then those segments will be provided to the user to choose from in an interactive gadget to keep only one AU per site after all buffering has finished. If the maximum buffer distance does not yield any stream segments, then those sites will be returned to the user for special QA and manual table completion after the report is finished. It is advised that you investigate the site further and potentially correct site coordinates in CEDS to prevent future problems. 

**Note: All stations you want to use in the assessment applications require AU information. If no AU information is provided to the app with a StationID, then the station will not appear in the app and could potentially be lost until further QA.**

The next chunk will bring in the riverine AU layer from the previous cycle, complete the snapping tool, ask for user input in the form of a gadget in the viewer panel if sites snap to multiple segments in a single buffer, and then organize results. You can see the progress and results of the tool below the chunk.

```{r riverineAUsnap, echo=FALSE}
# use a custom snapping tool to snap and organize 
riverineAUs <-  st_read('GIS/draft2018IR_AUs/va_2018_aus_riverine.shp') %>%
  st_transform(crs = 102003)  # transform to Albers for spatial intersection

Regional_Sites_AU <- snapAndOrganizeAU(conventionals_D_Region_NoAU, riverineAUs,  bufferDistances = seq(10,50,by=10))
# this buffers up to 100 m by 10 m
#Regional_Sites_AU <- snapAndOrganizeAU(conventionals_D_Region_NoAU, riverineAUs, bufferDistances = seq(10,100,by=10)) 
 # this buffers up to 50 m by 5 m
#Regional_Sites_AU <- snapAndOrganizeAU(conventionals_D_Region_NoAU, riverineAUs, bufferDistances = seq(10,50,by=5))

rm(riverineAUs) #clean up workspace
```

We can now filter the results to figure out sites that worked and sites that still need special attention. 

Below are the sites that still need work.
```{r conventionals_D_Region_NoAU still, echo=F}
conventionals_D_Region_NoAU <- filter(Regional_Sites_AU, is.na(ID305B_1))
DT::datatable(conventionals_D_Region_NoAU, escape=F, rownames = F, options = list(scrollX = TRUE))
#View(conventionals_D_Region_NoAU)
```

These sites will still be snapped to WQS, but they still need AU information before they can be run through the assessment applications. A reminder at the end of the report will encourage users to complete that step.


Below are the sites that worked. These can be combined with our conventionals_D_Region_AU layer now. 
```{r Regional_Sites_AUgood, echo=F}
Regional_Sites_AU <- filter(Regional_Sites_AU, !is.na(ID305B_1))
DT::datatable(Regional_Sites_AU, escape=F, rownames = F, options = list(scrollX = TRUE))
#View(Regional_Sites_AU)


conventionals_D_Region_AU <- rbind(conventionals_D_Region_AU, 
                                    Regional_Sites_AU %>% st_transform( st_crs(4326)), 
                                    conventionals_D_Region_NoAU %>% st_transform( st_crs(4326)))

rm(Regional_Sites_AU) # clean up workspace
```


### WQS Organization

Now all the stations need water quality standards attached as the final data organization step in order to be used by the assessment apps. Regardless of whether or not a site snapped to an AU, the report will walk users through connecting each site to WQS features. First, we will start with lakes to save the computationally heavier stream snapping tool for only the sites that need it.

#### Updated Reservoir WQS layer

We will bring in the updated WQS reservoir layer, to attach to stations that are in lacustrine AU's. 

```{r find Lacustrine AUs, echo=F}
# Bring reservoir AUs back in just for ID305B options
reservoirAUs <- as.character(st_read('GIS/draft2018IR_AUs/va_2018_aus_reservoir.shp')$ID305B)

# Separate lake stations from stream stations
reservoirStations <- filter(conventionals_D_Region_AU, ID305B_1 %in% unique(reservoirAUs) |
                              ID305B_2 %in% unique(reservoirAUs) | ID305B_3 %in% unique(reservoirAUs))

# Bring in reservoir WQS
reservoirWQS <- st_read('GIS/updatedWQS/reservoir.shp') %>%
   st_transform( st_crs(4326)) # transform to WQS84 for spatial intersection

# find lake WQS
reservoirStationsWQS <- st_join(reservoirStations, reservoirWQS, join = st_intersects) %>%
  filter(!is.na(OBJECTID)) # only keep rows where join successful
#View(reservoirStationsWQS)

# Run all stations that didn't connect to lake WQS polygon through streams snapping tool
riverineStationsWQS <- filter(conventionals_D_Region_AU, !(FDT_STA_ID %in% reservoirStationsWQS$FDT_STA_ID))
                              
# Make sure we didn't lose anyone
nrow(riverineStationsWQS) + nrow(reservoirStationsWQS) == nrow(conventionals_D_Region_AU)

rm(reservoirWQS);rm(reservoirAUs); rm(reservoirStations) # clean up workspace
```


All stations that did not attach to a lake AU will be run below as a stream site. This stream WQS snapping tool behaves the same way as the AU snapping tool; however, we are bringing in basins for snapping one at a time because the WQS files are so large. The tool will handle the organization and snapping by basin. You are only responsible for using the gadget that pops up in the Viewer panel when each basin prompts you. You can see the progress below the chunk. This is the most computationally heavy part of the process.


```{r attachStreamWQS, echo=F}
regionalOutput <- list()
# Loop through multiple basins
for (i in 1:length(unique(riverineStationsWQS$Basin))){
  Regional_Sites_AU_basin <- filter(riverineStationsWQS, Basin %in% unique(riverineStationsWQS$Basin)[i] )
  WQSsnaps <- snapAndOrganizeWQS(Regional_Sites_AU_basin, 'GIS/updatedWQS/', 
                                 basinNameSwitcher(unique(riverineStationsWQS$Basin)[i]),  seq(10,50,by=10))
  regionalOutput[[i]] <- WQSsnaps
}

# smash it all back together, change from lists to single dataframe
RegionalResults <- do.call("rbind", lapply(regionalOutput, data.frame))
#View(RegionalResults)
rm(WQSsnaps)# clean up workspace
```

Below is the table of riverine stations without WQS information attached. These will be returned to the user with a full list of stations that need individual attention. 

```{r RegionalResults table, echo=F}
RegionalResultsNoWQS <- filter(RegionalResults, is.na(OBJECTID))

DT::datatable(RegionalResultsNoWQS, escape=F, rownames = F, options = list(scrollX = TRUE))
```

Lastly, we need to combine the lake and stream stations into a single dataframe. After we combine we will separate out the stations that need either AU or WQS manual entry. We suggest using GIS to find this information. 

```{r combine WQS, echo=F}
# Reorder columns in reservoirStationsWQS to match RegionalResults to allow row binding
reservoirStationsWQS <- mutate(reservoirStationsWQS, Point.Unique.Identifier=NA, Buffer.Distance=NA, StreamType=NA) %>%
  select(FDT_STA_ID:COMMENTS, Point.Unique.Identifier, Buffer.Distance, OBJECTID:FCode, BASIN, WQS_COMMEN, 
         WATER_NAME, SEC, CLASS, SPSTDS, SECTION_DE, Basin_Code, PWS, Trout, Edit_Date, StreamType, 
         Tier_III, Backlog, Shape_Leng)%>%
  st_set_geometry(NULL)

RegionalResultsCombined <- rbind(RegionalResults,reservoirStationsWQS)
#View(RegionalResultsCombined)
```


This next chunk walks users through saving all the results. The results are saved to a default folder created in this directory called processedStationData. You can find the datasets you need to manually update before entering into the app there. Again, AU and WQS information as formatted by the above code is required for all sites you wish to analyze within the assessment apps. Failure to upload stations in the correct format ensures the app with either break or fail to analyze your data. It is advised that you fix the stations in the RegionalResultsCombined_REGIONCODE.csv. Other spreadsheets and shapefiles of stations missing information are provided for you to use in GIS to make corrections in the RegionalResultsComnined.csv.

```{r saveResults, echo=FALSE}
# This table has all unique stations within the selected region, regardless of AU or WQS connections.
write.csv(RegionalResultsCombined, paste('processedStationData/RegionalResultsCombined_',
                                         unique(assessmentLayerSelection$ASSESS_REG),'.csv',sep=''), row.names=FALSE)

# This table and shapefile highlights the stations that need AU information prior to use in assessment apps
RegionalResultsCombined_NoAU <- filter(RegionalResultsCombined, is.na(ID305B_1))
write.csv(RegionalResultsCombined_NoAU, paste('processedStationData/RegionalResultsCombined_NoAU_',
                                              unique(assessmentLayerSelection$ASSESS_REG),'.csv',sep=''), row.names=FALSE)
RegionalResultsCombined_NoAU_sf <- RegionalResultsCombined_NoAU %>%
  st_as_sf(coords = c("Longitude", "Latitude"),  # make spatial layer using these columns
           remove = F, # don't remove these lat/lon cols from df
           crs = 4326) # add coordinate reference system, needs to be geographic for now bc entering lat/lng,
st_write(RegionalResultsCombined_NoAU_sf, paste('processedStationData/RegionalResultsCombined_NoAU_',
         unique(assessmentLayerSelection$ASSESS_REG),'.shp',sep=''))

# This table and shapefile highlights the stations that need WQS information prior to use in assessment apps
RegionalResultsCombined_NoWQS <- filter(RegionalResultsCombined, is.na(OBJECTID))
write.csv(RegionalResultsCombined_NoWQS, paste('processedStationData/RegionalResultsCombined_NoWQS_',
                                              unique(assessmentLayerSelection$ASSESS_REG),'.csv',sep=''), row.names=FALSE)
RegionalResultsCombined_NoWQS_sf <- RegionalResultsCombined_NoWQS %>%
  st_as_sf(coords = c("Longitude", "Latitude"),  # make spatial layer using these columns
           remove = F, # don't remove these lat/lon cols from df
           crs = 4326) # add coordinate reference system, needs to be geographic for now bc entering lat/lng,
st_write(RegionalResultsCombined_NoWQS_sf, paste('processedStationData/RegionalResultsCombined_NoWQS_',
         unique(assessmentLayerSelection$ASSESS_REG),'.shp',sep=''))


# This table has all Estuary sites for your reference. None of the available assessment apps currently handle estuarine assessments
write.csv(conventionals_D_Region_Estuary, paste('processedStationData/EstuarineSites_',unique(assessmentLayerSelection$ASSESS_REG),'.csv',sep=''), row.names=FALSE)
```


## Split out riverine vs lake stations for assessment apps

Once you fix any stations that are missing AU or WQS information, you can read the entire dataset back to this script and have the program split the station tables into stations that will be analyzed by the lake assessment app and those that will be assessed by the rivers and streams assessment app. Packages and necessary datasets are reloaded in case you are not completing this step in the same session as the previous AU/WQS assignment session. 

** Minimal user input: if the file or file location you want to upload differs from the example given, you will need to update it to work with your configuration. **


```{r splitRiversAndLakes, echo=F}
#reload libraries in case you are starting from this point
library(tidyverse)
library(sf)

# Bring back in lake AUs from last cycle to filter input table
reservoirAUs <- st_read('GIS/draft2018IR_AUs/va_2018_aus_reservoir.shp') %>%
   st_transform( st_crs(4326)) # transform to WQS84 for spatial intersection


# This is the station table you have edited including all riverine and lake stations with all necessary AU and WQS information fixed for your region.
RegionalResultsCombinedAll <- read_csv('processedStationData/RegionalResultsCombined_BRRO_final.csv')

# Find all stations that have lake AUs
lakeStations <- filter(RegionalResultsCombinedAll, ID305B_1 %in% unique(reservoirAUs$ID305B) |
                         ID305B_2 %in% unique(reservoirAUs$ID305B) |
                         ID305B_3 %in% unique(reservoirAUs$ID305B))

# Make the rest riverine
riverStations <- filter(RegionalResultsCombinedAll, !(FDT_STA_ID %in% lakeStations)) 

# Save both for app use
write.csv(lakeStations, paste('processedStationData/RegionalResultsLake_',
                                         unique(assessmentLayerSelection$ASSESS_REG),'.csv',sep=''), row.names=FALSE)
write.csv(riverStations, paste('processedStationData/RegionalResultsRiverine_',
                                         unique(assessmentLayerSelection$ASSESS_REG),'.csv',sep=''), row.names=FALSE)
```



Contact Emma Jones (emma.jones@deq.virginia.gov) if you have any questions about this script.
